---
title: "Culcita_Project_All"
author: "Selma Mezger"
date: "15 2 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Import needed libraries
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(tidyr)

library(afex) # for GLMM
library(emmeans) # for posthoc analysis of GLMM

```



# Import and structuring of the datasets for the first 9 weeks (5 sites)

```{r}
Culcita <- read.csv("C://Users//Christian W//Documents//CushionStar_Matt//R predation raw 9w.csv", header = T, sep = ";")

colnames(Culcita)[1] <- "Week"
str(Culcita)

Culcita$Week <- as.factor(Culcita$Week)
Culcita$Site <- as.factor(Culcita$Site)
Culcita$Treatment <- as.factor(Culcita$Treatment)

str(Culcita)
```



# Plotting first 9 weeks

```{r}
# Summarize data to calculate mean and standard deviation per treatment for each week
summary_data_SE <- Culcita %>%
  group_by(Week, Treatment) %>%
  summarize(
    Recruits_Mean = mean(Recruits.Predation),
    Recruits_SE = sd(Recruits.Predation) / sqrt(n()), # Calculate standard error
    Transplants_Mean = mean(Transplants.Predation),
    Transplants_SE = sd(Transplants.Predation) / sqrt(n()) # Calculate standard error
  )


# Plot with standard error
# Create a line plot with summarized data
Plot9 <- ggplot(data = summary_data_SE, aes(x = Week)) +
  # Add line and points for Recruits.Predation
  geom_line(aes(y = Recruits_Mean, linetype = Treatment, group = paste(Treatment, "Recruits"), color = "Recruits"), size = 1, alpha = 0.8) +
  geom_point(aes(y = Recruits_Mean, color = "Recruits"), size = 3, alpha = 0.8) +
  # Add error bars for Recruits.Predation
  geom_errorbar(aes(ymin = Recruits_Mean - Recruits_SE, ymax = Recruits_Mean + Recruits_SE, color = "Recruits"), width = 0.2, alpha = 0.8) +
  # Add line and points for Transplants.Predation
  geom_line(aes(y = Transplants_Mean, linetype = Treatment, group = paste(Treatment, "Transplants"), color = "Transplants"), size = 1, alpha = 0.8) +
  geom_point(aes(y = Transplants_Mean, color = "Transplants"), size = 3, alpha = 0.8) +
  # Add error bars for Transplants.Predation
  geom_errorbar(aes(ymin = Transplants_Mean - Transplants_SE, ymax = Transplants_Mean + Transplants_SE, color = "Transplants"), width = 0.2, alpha = 0.8) +
  # Adjust legend and axis labels
  labs(x = "Week", y = expression(atop("Accumulated", paste(italic("C. schemideliana")*" Predation (%)")))) +
  scale_color_manual(values = c("Recruits" = "#21908CFF", "Transplants" = "#440154FF"), 
                     labels = c("Recruits", "Transplants"), 
                     name = "Variables") +
  scale_linetype_manual(values = c("Control" = "solid", "Test" = "dashed"), labels = c("Control", "Test")) +
  theme_bw() +
  # Adjust linetype scale
  theme(legend.key.size = unit(1.0, "cm"),
        legend.key.height = unit(0.5, "cm"),  # smaller dashed linetype
        panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        axis.text = element_text(size = 12),   # Increase size of axis labels
        axis.title = element_text(size = 13),  # Increase size of axis titles
        axis.ticks = element_line(size = 0.5)) # Increase size of axis ticks

Plot9

ggsave("C://Users//Christian W//Documents//CushionStar_Matt//Figure1.png", Plot9, width = 7.8, height = 5, dpi = 600)

```


# Plotting 27 weeks

```{r}
# Upload data for plotting
Culcita27 <- read.csv("C://Users//Christian W//Documents//CushionStar_Matt//R predation outplants raw 27w.csv", header = T, sep = ";")

colnames(Culcita27)[1] <- "Week"
str(Culcita27)
# n this case we don't turn time into a factor, so it stays a continuous variable for the plotting along the x axis
Culcita27$Site <- as.factor(Culcita27$Site)
Culcita27$Treatment <- as.factor(Culcita27$Treatment)

str(Culcita27)

# Summarize the data to calculate means and standard errors
summary_data_Culcita27 <- Culcita27 %>%
  group_by(Week, Treatment) %>%
  summarize(
    Culcita_Mean = mean(Culcita.Predation),
    Culcita_SE = sd(Culcita.Predation) / sqrt(n())
  )


# Create a line plot with summarized data
Plot27 <- ggplot(data = summary_data_Culcita27, aes(x = Week)) +
  # Add line and points for Culcita Predation
  geom_line(aes(y = Culcita_Mean, linetype = Treatment, group = Treatment, color = Treatment), size = 1, alpha = 0.8) +
  geom_point(aes(y = Culcita_Mean, color = Treatment), size = 3, alpha = 0.8) +
  # Add error bars for Culcita Predation
  geom_errorbar(aes(ymin = Culcita_Mean - Culcita_SE, ymax = Culcita_Mean + Culcita_SE, color = Treatment), width = 0.2) +
  # Add vertical line at week 9
  geom_vline(xintercept = 9, linetype = "dashed", color = "black") +
  # Adjust legend and axis labels
    labs(x = "Week", y = expression(atop("Accumulated", paste(italic("C. schemideliana")*" Predation (%)")))) +
  scale_linetype_manual(values = c("Control" = "solid", "Test" = "dashed"), labels = c("Control", "Test")) +
  scale_color_manual(values = c("#3B528BFF", "#5DC863FF"), labels = c("Control", "Test")) + # Color assignment
  scale_x_continuous(breaks = seq(1, max(summary_data_Culcita27$Week), by = 2)) + # Adjust x-axis breaks
  theme_bw() +
  theme(legend.key.size = unit(1.0, "cm"),
        legend.key.height = unit(0.5, "cm"),  
        panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor.x = element_blank(),  # Remove vertical minor gridlines
        axis.text = element_text(size = 12),   
        axis.title = element_text(size = 13),  
        axis.ticks = element_line(size = 0.5))

Plot27

ggsave("C://Users//Christian W//Documents//CushionStar_Matt//Figure2.png", Plot27, width = 7.8, height = 5, dpi = 600)

```






# Compute Linear Mixed model for the first 9 weeks of experiment
```{r}

## Recruits
# Build the GLMM and obtain the summary of the model
Recruits_New<- mixed(Recruits.Predation~factor(Treatment)*factor(Week)+(1|Site),data=Culcita,check.contrasts=FALSE) # here we have treatment as the fixed effect, and allow random intercepts per site and week
summary(Recruits_New)
Recruits_New

# Obtain the estimated marginal means
rec_emmeans_model <- emmeans(Recruits_New, specs = ~ Treatment | Week)

# Display the results
summary(rec_emmeans_model, type = "response")


# Perform pairwise comparisons with Tukey adjustment
rec_pairwise_comp <- emmeans(Recruits_New, pairwise ~ Treatment | Week, adjust = "tukey")

# Display the results
summary(rec_pairwise_comp)




## Transplants
# Build the GLMM and obtain the summary of the model
Transplant_New<- mixed(Transplants.Predation~factor(Treatment)*factor(Week)+(1|Site),data=Culcita,check.contrasts=FALSE) # here we have treatment as the fixed effect, and allow random intercepts per site and week
summary(Transplant_New)
Transplant_New

# Obtain the estimated marginal means
tran_emmeans_model <- emmeans(Transplant_New, specs = ~ Treatment | Week)

# Display the results
summary(tran_emmeans_model, type = "response")


# Perform pairwise comparisons with Tukey adjustment
tran_pairwise_comp <- emmeans(Transplant_New, pairwise ~ Treatment | Week, adjust = "tukey")

# Display the results
summary(tran_pairwise_comp)




# Comparison transplants vs recruits predation

# Upload new dataframe
CulcitaTransRecr <- read.csv("C://Users//Christian W//Documents//CushionStar_Matt//R predation raw 9w RecruitTransplantComp.csv", header = T, sep = ";")


colnames(CulcitaTransRecr)[1] <- "Week"
str(CulcitaTransRecr)

CulcitaTransRecr$Week <- as.factor(CulcitaTransRecr$Week)
CulcitaTransRecr$Site <- as.factor(CulcitaTransRecr$Site)
CulcitaTransRecr$Treatment <- as.factor(CulcitaTransRecr$Treatment)


str(CulcitaTransRecr)

CulcitaTransRecrCntr <- subset(CulcitaTransRecr, Treatment == "Control")

# Run linear mixed model
predation_model <- mixed(Predation_Amount ~ Predation_Type * Week + (1|Site),
                         data = CulcitaTransRecrCntr,
                         check.contrasts = FALSE)

# Summary of the model
summary(predation_model)
predation_model

# Obtain the estimated marginal means
emmeans_model <- emmeans(predation_model, specs = ~ Predation_Type | Week)

# Display the results
summary(emmeans_model, type = "response")

# Perform pairwise comparisons with Tukey adjustment
pairwise_comp <- emmeans(predation_model, pairwise ~ Predation_Type | Week, adjust = "tukey")

# Display the results
summary(pairwise_comp)





# Comparison Acro vs Poci recruits predation

# Upload new dataframe
Recruits9W <- read.csv("C://Users//Christian W//Documents//CushionStar_Matt//9 weeks recruits by quadrat.csv", header = T, sep = ";")


colnames(Recruits9W)[1] <- "Week"
str(Recruits9W)

Recruits9W$Week <- as.factor(Recruits9W$Week)
Recruits9W$Site <- as.factor(Recruits9W$Site)
Recruits9W$Treatment <- as.factor(Recruits9W$Treatment)
Recruits9W$Genus <- as.factor(Recruits9W$Genus)


str(Recruits9W)

RecruitsCntr <- subset(Recruits9W, Treatment == "Control")
RecruitsTest <- subset(Recruits9W, Treatment == "Test")

Recruits9W %>%
  group_by(Week, Treatment, Genus) %>%
  get_summary_stats(Culcita.Predation, type = "mean_sd")


# Control Site
# Run linear mixed model
RecruitCntr_model <- mixed(Culcita.Predation ~ Genus * Week + (1|Site),
                         data = RecruitsCntr,
                         check.contrasts = FALSE)

# Summary of the model
summary(RecruitCntr_model)
RecruitCntr_model



# Test Site
# Run linear mixed model
RecruitTest_model <- mixed(Culcita.Predation ~ Genus * Week + (1|Site),
                         data = RecruitsTest,
                         check.contrasts = FALSE)

# Summary of the model
summary(RecruitTest_model)
RecruitTest_model






# Comparison Acro vs Poci transplants predation

# Upload new dataframe
Transplants9W <- read.csv("C://Users//Christian W//Documents//CushionStar_Matt//9 weeks outplants by cluster.csv", header = T, sep = ";")


colnames(Transplants9W)[1] <- "Week"
str(Transplants9W)

Transplants9W$Week <- as.factor(Transplants9W$Week)
Transplants9W$Site <- as.factor(Transplants9W$Site)
Transplants9W$Designation <- as.factor(Transplants9W$Designation)
Transplants9W$Genus <- as.factor(Transplants9W$Genus)


str(Transplants9W)

TransplantsCntr <- subset(Transplants9W, Designation == "Control")
TransplantsTest <- subset(Transplants9W, Designation == "Test")

Transplants9W %>%
  group_by(Week, Designation, Genus) %>%
  get_summary_stats(Culcita.Predation, type = "mean_sd")


# Control Site
# Run linear mixed model
TransplantsCntr_model <- mixed(Culcita.Predation ~ Genus * Week + (1|Site),
                         data = TransplantsCntr,
                         check.contrasts = FALSE)

# Summary of the model
summary(TransplantsCntr_model)
TransplantsCntr_model

# Obtain the estimated marginal means
emmeans_model <- emmeans(TransplantsCntr_model, specs = ~ Genus | Week)

# Display the results
summary(emmeans_model, type = "response")

# Perform pairwise comparisons with Tukey adjustment
pairwise_comp <- emmeans(TransplantsCntr_model, pairwise ~ Genus | Week, adjust = "tukey")

# Display the results
summary(pairwise_comp)



# Test Site
# Run linear mixed model
TransplantsTest_model <- mixed(Culcita.Predation ~ Genus * Week + (1|Site),
                         data = TransplantsTest,
                         check.contrasts = FALSE)

# Summary of the model
summary(TransplantsTest_model)
TransplantsTest_model

```








# 27 weeks test vs control comparison

# Import and structuring of the datasets for the recovery period (Week 9-27)

```{r}
# first I input the subset of the experimental data
Culcita27_short <- Culcita27[Culcita27$Week %in% c(9, 15, 21, 27), ]

str(Culcita27_short)

```



# GLMM for the recovery phase

```{r}
# Build the GLMM and obtain the summary of the model
Culcita27_short_model<- mixed(Culcita.Predation~factor(Treatment)*factor(Week)+(1|Site),data=Culcita27_short,check.contrasts=FALSE) # here we have treatment as the fixed effect, and allow random intercepts per site and week
summary(Culcita27_short_model)
Culcita27_short_model

# Obtain the estimated marginal means
Culcita27_short_emmeans_model <- emmeans(Culcita27_short_model, specs = ~ Treatment | Week)

# Display the results
summary(Culcita27_short_emmeans_model, type = "response")


# Perform pairwise comparisons with Tukey adjustment
Culcita27_short_pairwise_comp <- emmeans(Culcita27_short_model, pairwise ~ Treatment | Week, adjust = "tukey")

# Display the results
summary(Culcita27_short_pairwise_comp)
```


# quickly double checking the assumption that the slopes for both curves are the same using a simple linear regression model

```{r}
# Fit a linear model with interaction term
model_short <- lm(Culcita.Predation ~ Week * Treatment, data = Culcita27_short)

# Print the summary of the model
summary(model_short)
```





# Culcita density

```{r}
CulcitaDensity <- read.csv("C://Users//Christian W//Documents//CushionStar_Matt//Density.csv", header = T, sep = ",")

colnames(CulcitaDensity)[1] <- "Site"
str(CulcitaDensity)

CulcitaDensity$Site <- as.factor(CulcitaDensity$Site)

str(CulcitaDensity)


# test for normality
# Build the linear model
model  <- lm(Density ~ Site, data = CulcitaDensity)
# Create a QQ plot of residuals
ggqqplot(residuals(model))
# Compute Shapiro-Wilk test of normality
shapiro_test(residuals(model))

CulcitaDensity %>%
  group_by(Site) %>%
  shapiro_test(Density)


# test homogeneity
CulcitaDensity %>% levene_test(Density ~ Site)

# data is neither normal nor homogenous --> non-parametric test

res.kruskal <- CulcitaDensity %>% kruskal_test(Density ~ Site)
res.kruskal # p = 0.00318

CulcitaDensity %>% kruskal_effsize(Density ~ Site) # large, eta2[H] = 0.22

pwc <- CulcitaDensity %>% dunn_test(Density ~ Site, p.adjust.method = "bonferroni")
pwc # WS to OOTB R (0.014) and WS to OOTS L (0.009)



# Plotting
library(viridis)

# Define the colors for different sites using the viridis color palette
site_colors <- viridis_pal()(n_distinct(CulcitaDensity$Site))


# Create the boxplot with original measurements as dots
p <- ggplot(CulcitaDensity, aes(x = Site, y = Density, fill = Site)) +
  geom_boxplot(alpha = 0.7) +  # Add boxplot with transparency
  geom_point(aes(color = Site), position = position_jitter(width = 0.2), size = 2) +  # Add points with jitter for better visibility
  scale_fill_manual(values = site_colors) +  # Set colors for boxplot
  scale_color_manual(values = site_colors) +  # Set colors for points
  labs(x = "Site", y = expression(italic("C. schemideliana")*" density/"~100 * m^2)) +
  ylim(0, 8) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),  # Remove vertical gridlines
        panel.grid.minor.x = element_blank(),  # Remove vertical minor gridlines
        axis.text = element_text(size = 12),   
        axis.title = element_text(size = 13),  
        axis.ticks = element_line(size = 0.5)) +
  # Use minimal theme
  theme(legend.position = "none")  # Remove legend

# Calculate maximum density for each site
max_density <- aggregate(Density ~ Site, data = CulcitaDensity, FUN = max)

# Add annotation letters
annotation_df <- data.frame(
  Site = max_density$Site,
  Max_Density = max_density$Density,
  Letter = c("ab", "ab", "a", "a", "b")
)

DensityPlot <- p + 
  geom_text(data = annotation_df, aes(x = Site, y = Max_Density, label = Letter), vjust = -1, size = 6)

DensityPlot

ggsave("C://Users//Christian W//Documents//CushionStar_Matt//Figure3.png", DensityPlot, width = 7.8, height = 5, dpi = 600)
```

